---
title: "fquantdr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fquantdr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

# Getting Started

To get started you need to first install the package using the command
```{r, eval = FALSE}
install.packages("fquantdr")
```

and then use it during any 'R' session by issuing the command
```{r setup}
library(fquantdr)
```

For an overview of the available help files of the package use
```{r, eval = FALSE}
help(package = "fquantdr")
```

The package includes the following functions (Keep updated):

* `fundata` Generate functional data based on Fourier basis function
* `fcqs`
* `sonf`

<!-- * `discrete` Discretize conversion of a continuous numeric vector -->
<!-- * `gramatrix` Computes the center gram matrix -->
<!-- * `matpower` Power of a matrix -->
<!-- * `mcorr` Multiple correlation between two random vectors -->
<!-- * `mppower` Moore-Penrose type power -->
<!-- * `qmat` Computes a transformed identity matrix -->
<!-- * `rigpower` Ridge-regression type power -->
<!-- * `slav` Compute slice average  -->
<!-- * `slprob` Compute slice proportions  -->
<!-- * `symmetry` Symmtrize Matrix -->

To get help on a specific function type `help(function.name)` or `?function.name` for a convenient shorthand. Try
```{r, eval = FALSE}
help(fcqs)
?cqs
```

For further examples, refer to the coding examples listed in the documentation for each `fquantdr` function.

# Functional central quantile subspace
## Overview

The overall goal of the `fcqs` function is to identify a finite linear operator $L_\tau(X)$, where $L_{\tau} \in \mathscr{B}(\mathscr{H}, \mathbb{R}^{d_\tau})$ and $d_{\tau} \in \mathbb{N}$, such that the functional central quantile subspace is $\overline{\mathrm{ran}}(L_\tau^*)$. This allows us to replace the random function $X$, defined as $X = (X^1, \dots, X^p): \Omega \rightarrow \bigoplus_{i=1}^{p} \mathcal{H}_i$, with the lower-dimensional linear operator $L_\tau(X)$. To achieve this, Christou et al. (2024+) proved two important results:

* $\beta_\tau^* \in \mathcal{S}_{Q_{\tau}(Y|X)}$, where $\beta_\tau^*$ is the slope vector from regressing the conditional quantile on $X$, i.e.,

\begin{eqnarray*}
(\alpha_{\tau}^*, \beta_{\tau}^*) = \arg \min_{(a_{\tau}, b_{\tau})} E [Q_{\tau}\{Y|L(X)\} - a_{\tau} - \langle b_{\tau}, X \rangle_{\mathscr{H}}]^2
\end{eqnarray*}
where $L \in \mathscr{B}(\mathscr{H}, \mathbb{R}^d)$, such that the functional central subspace is $\overline{\mathrm{ran}}(L^*)$ (Li and Song, 2022). This implies an initial dimension reduction using $L(X)$, making the nonparametric estimation of $Q_\tau(Y|L(X))$ tractable. 

* $E({\langle b_\tau, X \rangle}_\mathscr{H} X ) \in \Sigma_{XX} \mathcal{S}_{Q_\tau (Y|X)}$, where $b_\tau \in \mathcal{S}_{Q_\tau (Y|X)}$

The above two results imply the following:

1. If the dimension of the $\tau$-FCQS is known to be one, then we can fit a linear quantile regression model of $Q_\tau\{Y|L(X)\}$ on $X$ and report the 
slope vector $\beta_\tau^*$ as the basis vector (double check it).

1. If the dimension of the $\tau$-FCQS is known to be greater than one, then we can set $b_{\tau 0} = \beta_\tau^*$ as the initial vector and then create more vectors using $b_{\tau j} = E(\langle b_{\tau j-1}, X \rangle_\mathscr{H} X)$ for $j = 1, ..., m$, where $m = pq-1$ (double check). In order to obtain more linear independent vector, we can let $B_\tau = \sum_{j=0}^m b_{\tau j} \otimes b_{\tau j}$, then perform the generalized eigenvalue problem $B_\tau v_{\tau j} = \lambda \Sigma_{XX} v_{\tau j}$ and choose the $v_{\tau 1}, ..., v_{\tau d_\tau}$ be the set of solutions. 

All of the above are incorporated in the `fcqs` function.

## The algorithm at a glance

The first step of the algorithm requires to fit a linear regression model of $Q_\tau \{Y|L(X) \}$ on $L(X)$ (or X?), which implies the non-parametric estimation of the conditional quantile. The second step of the algorithm relies on estimating an expectation and performing a generalized eigenvalue problem.  Specifically,

1. For $u=1, \dots, n$, let $Y_u$ be an independent and identically distributed (iid) sample from $Y$, and let $X_1, \dots, X_n$ be an independent sample from the random element $X = (X^1, \dots, X^p)$, such that $X_u = (X_u^1, \dots, X_u^p)^{\top}$.

1. We apply a standard dimension reduction technique to replace \( X \) with \( L(X) \). Specifically, we use Functional Sliced Inverse Regression (FSIR) as proposed by Ferré and Yao (2003). Using the `mfsir` function of the presented package, we estimate the \( d \)-dimensional predictors \(\widehat{L}(X_u)\), for \( u = 1, \dots, n \)

1. For each $u=1, \dots, n$, we use the local linear quantile estimation method of Guerre and Sabbah (2012) to estimate $\widehat{Q}_{\tau}\{Y| \widehat{L}(X_u)\}$

1. We set $\widehat{\beta}_{\tau}$ to be 
\begin{eqnarray*}
(\widehat{\alpha}_{\tau}, \widehat{\beta}_{\tau}) = \arg \min_{(a_{\tau}, b_{\tau})} \sum_{u=1}^{n} [\widehat{Q}_{\tau}\{Y | \widehat{L}(X_u)\} - a_{\tau} - \langle b_{\tau}, X_u \rangle_{\mathscr{H}}]^2,
\end{eqnarray*}
For this we use the `sonf` function of the presented package.

1. We set $\widehat{\beta}_{\tau 0} = \widehat{\beta}_{\tau}$, and, for $j=1, \dots, m$, we form
\begin{eqnarray*}
\widehat{\beta}_{\tau j} = n^{-1} \sum_{u=1}^{n}\langle \widehat{\beta}_{\tau j-1}, X_u \rangle_{\mathscr{H}}X_u.
\end{eqnarray*}

1. After obtaining $\widehat{\beta}_{\tau 0}, \widehat{\beta}_{\tau 1}, \dots, \widehat{\beta}_{\tau m}$, we set $\widehat{B}_{\tau} = \sum_{j=0}^{m} \widehat{\beta}_{\tau j} \otimes \widehat{\beta}_{\tau j}$ and focus on the eigenvalue problem (should we include more details in this step?)

## Ouput of `fcqs`


