---
title: "fquantdr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fquantdr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

# Getting Started

To get started you need to first install the package using the command
```{r, eval = FALSE}
install.packages("fquantdr")
```

and then use it during any 'R' session by issuing the command
```{r setup}
library(fquantdr)
```

For an overview of the available help files of the package use
```{r, eval = FALSE}
help(package = "fquantdr")
```

The package includes the following functions (Keep updated):

* 'fundata' Generate functional data based on Fourier basis function
* 'fcqs'
* 'sonf'
<!-- * 'discrete' Discretize conversion of a continuous numeric vector -->
<!-- * 'gramatrix' Computes the center gram matrix -->
<!-- * 'matpower' Power of a matrix -->
<!-- * 'mcorr' Multiple correlation between two random vectors -->
<!-- * 'mppower' Moore-Penrose type power -->
<!-- * 'qmat' Computes a transformed identity matrix -->
<!-- * 'rigpower' Ridge-regression type power -->
<!-- * 'slav' Compute slice average  -->
<!-- * 'slprob' Compute slice proportions  -->
<!-- * 'symmetry' Symmtrize Matrix -->

To get help on a specific function type 'help(function.name)' or '?function.name' for a convenient shorthand. Try
```{r, eval = FALSE}
help(fcqs)
?cqs
```

For further examples, refer to the coding examples listed in the documentation for each 'fquantdr' function.
