---
title: "Dimension Reduction Techniques for Conditional Quantiles for Functional Data in R: A Vignette"
author:  
- name: Shanshan Wang, Eliana Christou
  affiliation: University of North Carolina at Charlotte
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dimension Reduction Techniques for Conditional Quantiles for Functional Data in R: A Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
abstract: "Functional data analysis has received increasing attention in recent years due to its ability to handle infinite-dimensional functional predictors, which require dimension reduction techniques. However, existing research primarily relies on traditional mean regression methods, with limited focus on quantile regression. `fquantdr` is an `R` package that addresses this gap by performing dimension reduction for conditional quantiles in functional data, identifying the directions that span the Functional Central Quantile Subspace (FCQS)."
keywords: "functional central quantile subspace, dimension reduction, conditional quantiles, functional data analysis"

---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

Functional data analysis (FDA) is a field of statistics that considers data as realizations of random functions. It was first formally introduced by Ramsay and Dalzell (1991). The challenge with functional data lies in their *infinite-dimensional* space, which requires the use of dimension reduction techniques to replace them with finite data. For references, see Ferré and Yao (2003), Wang et al. (2013), Lian and Li (2014), and Li and Song (2022). Many functional data sets exhibit characteristics such as skewness and outliers, which make quantile regression (QR) a more suitable approach in these cases.

Quantile regression (QR) has gained significant attention since its inception by Koenker and Bassett (1978), becoming a valuable tool in research areas where non-central parts of the data are important. Despite its extensive use, QR has seen limited application in the context of functional data. For reference, see Cardot et al. (2005), Kato (2012), and Chen and Müller (2012).


Recently, Christou et al. (2024+) proposed a novel technique to identify a finite linear operator that captures all the information about the conditional quantile for functional data. Specifically, for a univariate response $Y$ and a random element $X = (X^1, \ldots, X^p)$ in a Hilbert space $\mathscr{H}$, the method aims to find a finite linear operator $L_\tau \in \mathscr{B}(\mathscr{H}, \mathbb{R}^{d_\tau})$ for $\tau \in (0, 1)$ and $d_\tau \leq p$, such that $Y$ and $Q_\tau(Y|X)$ are conditionally independent given $L_\tau(X)$. Here, $Q_\tau(Y|X)$ denotes the $\tau$th conditional quantile of $Y$ given $X$. The space $\overline{\text{ran}}(L_\tau^*)$ is referred to as the *$\tau$th functional quantile dimension reduction subspace* for the regression of $Y$ on $X$. The intersection of all $\tau$th dimension reduction subspaces is called the *$\tau$th functional central quantile subspace* ($\tau$-FCQS) and is denoted by $\mathcal{S}_{Q_\tau(Y|X)}$.

The purpose of this vignette is to demonstrate the implementation of the functions available in the fquantdr package through various examples. The main function of the package `fcqs` returns the estimated directions of the $\tau$-FCQS.

# Getting Started

To get started you need to first install the package using the command
```{r, eval = FALSE}
install.packages("fquantdr")
```

After installation, you can load the package in any R session by issuing the command:
```{r setup}
library(fquantdr)
```

For an overview of the available help files of the package use
```{r, eval = FALSE}
help(package = "fquantdr")
```

The package includes the following functions:

* `fcqs` Functional central quantile subspace
* `fundata` Functional data
* `mcorr` Multiple correlation
* `mfsir` Functional Sliced Inverse Regression
* `sonf` Scalar-on-function regression

<!-- * `discrete` Discretize conversion of a continuous numeric vector -->
<!-- * `gramatrix` Computes the center gram matrix -->
<!-- * `matpower` Power of a matrix -->
<!-- * `mppower` Moore-Penrose type power -->
<!-- * `qmat` Computes a transformed identity matrix -->
<!-- * `rigpower` Ridge-regression type power -->
<!-- * `slav` Compute slice average  -->
<!-- * `slprob` Compute slice proportions  -->
<!-- * `symmetry` Symmtrize Matrix -->

You can find detailed descriptions and examples for each function by using the `help()` function or the `?` operator in R, specifying the function name. For example, to get help on the `fcqs` function, use:
```{r, eval = FALSE}
help(fcqs)
?fcqs
```

For further examples, refer to the coding examples listed in the documentation for each `fquantdr` function.

# Functional central quantile subspace
## Overview

The overall goal of the `fcqs` function is to identify a finite linear operator $L_\tau(X)$, where $L_{\tau} \in \mathscr{B}(\mathscr{H}, \mathbb{R}^{d_\tau})$ and $d_{\tau} \in \mathbb{N}$, such that the functional central quantile subspace is $\overline{\mathrm{ran}}(L_\tau^*)$. This allows us to replace the random function $X$, defined as $X = (X^1, \dots, X^p): \Omega \rightarrow \bigoplus_{i=1}^{p} \mathcal{H}_i$, with the lower-dimensional linear operator $L_\tau(X)$. To achieve this, Christou et al. (2024+) proved two important results:

* $\beta_\tau^* \in \mathcal{S}_{Q_{\tau}(Y|X)}$, where $\beta_\tau^*$ is the slope vector from regressing the conditional quantile on $X$, i.e.,

\begin{eqnarray*}
(\alpha_{\tau}^*, \beta_{\tau}^*) = \arg \min_{(a_{\tau}, b_{\tau})} E [Q_{\tau}\{Y|L(X)\} - a_{\tau} - \langle b_{\tau}, X \rangle_{\mathscr{H}}]^2
\end{eqnarray*}
where $L \in \mathscr{B}(\mathscr{H}, \mathbb{R}^d)$, such that the functional central subspace is $\overline{\mathrm{ran}}(L^*)$ (Li and Song, 2022). This implies an initial dimension reduction using $L(X)$, making the nonparametric estimation of $Q_\tau(Y|L(X))$ tractable. 

* $E({\langle b_\tau, X \rangle}_\mathscr{H} X ) \in \Sigma_{XX} \mathcal{S}_{Q_\tau (Y|X)}$, where $b_\tau \in \mathcal{S}_{Q_\tau (Y|X)}$

The above two results imply the following:

First, we set $b_{\tau 0} = \beta_\tau^*$ as the initial vector and then create more vectors using $b_{\tau j} = E(\langle b_{\tau j-1}, X \rangle_\mathscr{H} X)$ for $j = 1, ..., m$, where $m = pq-1$, $p$ is the number of predictor variables and $q$ is the number of basis functions used to approximate the functional predictors. In order to obtain more linear independent vector, we can let $B_\tau = \sum_{j=0}^m b_{\tau j} \otimes b_{\tau j}$, then perform the generalized eigenvalue problem $B_\tau v_{\tau j} = \lambda \Sigma_{XX} v_{\tau j}$ and choose the $v_{\tau 1}, ..., v_{\tau d_\tau}$ be the set of solutions. 

All of the above are incorporated in the `fcqs` function.

## The algorithm at a glance

The first step of the algorithm requires to fit a linear regression model of $Q_\tau \{Y|L(X) \}$ on $X$, which implies the non-parametric estimation of the conditional quantile. The second step of the algorithm relies on estimating an expectation and performing a generalized eigenvalue problem.  Specifically,

1. For $u=1, \dots, n$, let $Y_u$ be an independent and identically distributed (iid) sample from $Y$, and let $X_1, \dots, X_n$ be an independent sample from the random element $X = (X^1, \dots, X^p)$, such that $X_u = (X_u^1, \dots, X_u^p)^{\top}$.

1. We apply a standard dimension reduction technique to replace \( X \) with \( L(X) \). Specifically, we use Functional Sliced Inverse Regression (FSIR) as proposed by Ferré and Yao (2003). Using the `mfsir` function of the presented package, we estimate the \( d \)-dimensional predictors \(\widehat{L}(X_u)\), for \( u = 1, \dots, n \)

1. For each $u=1, \dots, n$, we use the local linear quantile estimation method of Guerre and Sabbah (2012) to estimate $\widehat{Q}_{\tau}\{Y| \widehat{L}(X_u)\}$.
For this we use `llqr` function of the `quantdr` package

1. We set $\widehat{\beta}_{\tau}$ to be 
\begin{eqnarray*}
(\widehat{\alpha}_{\tau}, \widehat{\beta}_{\tau}) = \arg \min_{(a_{\tau}, b_{\tau})} \sum_{u=1}^{n} [\widehat{Q}_{\tau}\{Y | \widehat{L}(X_u)\} - a_{\tau} - \langle b_{\tau}, X_u \rangle_{\mathscr{H}}]^2,
\end{eqnarray*}
For this we use the `sonf` function of the presented package.

1. We set $\widehat{\beta}_{\tau 0} = \widehat{\beta}_{\tau}$, and, for $j=1, \dots, m$, we form
\begin{eqnarray*}
\widehat{\beta}_{\tau j} = n^{-1} \sum_{u=1}^{n}\langle \widehat{\beta}_{\tau j-1}, X_u \rangle_{\mathscr{H}}X_u.
\end{eqnarray*}

1. After obtaining $\widehat{\beta}_{\tau 0}, \widehat{\beta}_{\tau 1}, \dots, \widehat{\beta}_{\tau m}$, we set $\widehat{B}_{\tau} = \sum_{j=0}^{m} \widehat{\beta}_{\tau j} \otimes \widehat{\beta}_{\tau j}$ and focus on the eigenvalue problem and solving $\Sigma_{XX}^{\dagger 1/2} B_{\tau} \Sigma_{XX}^{\dagger 1/2} \eta_{\tau} = \lambda \eta_{\tau}, \quad v_{\tau} = \Sigma_{XX}^{\dagger 1/2} \eta_{\tau}$.

## Ouput of `fcqs`
In this section, we begin by simulating data from the homoscedastic single-index model.
\begin{eqnarray*}
\quad Y &= \frac{1}{0.5 + (\langle \beta_1, X \rangle_{\mathscr{H}} +1)^2} + 0.2\varepsilon, \ \text{where} \ \mathcal{S}_{Q_{\tau}(Y|X)} = \mathrm{span} \{\beta_1\}
\end{eqnarray*}
where \(\beta_1(t)=\psi_1(t)\) is the first eigenfunction of \(\Sigma_{XX}\). The predictor \(X\) is generated using the Karhunen-Loève expansion. The error term \(\varepsilon\) is generated from three different distributions: (1) standard normal \(\mathcal{N}(0, 1)\), (2) chi-square with three degrees of freedom \(\mathcal{X}_3^2\), and (3) exponential with rate of 0.5 (Exp(0.5)) distribution. We focus on the estimation of the 0.5-FCQS
```{r}
set.seed(1234)
# Parameters for generating functional data
n <- 100
p <- 5
nbasis <- 4
tau <- 0.5
d_tau <- 1
d_DR <- 1
time <- seq(0, 1, length.out = 101)
eta <- matrix(stats::rnorm(n * p * nbasis), nrow = n, ncol = p * nbasis)

# Generate functional data according to KL expanasion
result <- fundata(n, p, nbasis, time, eta)
xc <- result$xc
P <- eigen(cov(eta))$vectors
Q <- diag(eigen(cov(eta))$values)

# this is the inner products <b1, X>, <b2, X>, ... 
mfpca.scores <- eta %*% P

# Generate error for data
error <- rnorm(n)

# Generate y
y <- 1 / (0.5 + (mfpca.scores[, 1] + 1)^2) + 0.2 * error
```
The `fcqs` function operates by specifying a three-dimensional $n \times nt \times p$ array $X$, a vector of the response variable $Y$, a vector of time points $nt$ at which the functional data is evaluated, the number of basis functions for smoothing the functional data, and a quantile level $\tau$. The dimension of the $\tau$-FCQS is an integer between 1 and $p$. For this example, $d_{\tau} = 1$ for every quantile level $\tau$.

We now discuss the output of the `fcqs` function

```{r}
output1 <- fcqs(xc, y, time, nbasis, tau, d_tau)
output1
```
The output of the `fcqs` function includes two primary components:

1. `ffun`: The functional parameters that span the FCQS. These are the estimated directions in the functional central quantile subspace, which reflect the significant functional predictors for the quantile regression.

1. `betax`: The resulting sufficient predictor, calculated as the inner product between `ffun` and `x`. This represents the reduced-dimensional representation of the functional data that retains the essential information for the quantile regression.

If we want to measure the multiple correlation between the true and estimated predictors, we can use the `mcorr` function. Specifically, for random vectors $U$ and $V$ of the same dimension $d$, we define $C_{UU}, C_{UV},$ and $C_{VV}$ as the sample covariance matrices.  Then, the multiple correlation between $U$ and $V$ is given by:
\begin{eqnarray*}
mcorr(U, V) = tr(C_{VV}^{-1/2} C_{VU} C_{UU}^{-1} C_{UV} C_{VV}^{-1/2}),
\end{eqnarray*}

```{r}
# define true inner products for FCQS
uu_fcqs <- mfpca.scores[, 1:d_tau]

# Compute the multiple correlation between the true and estimated predictors
mcorr.fcqs.vec <- mcorr(output1$betax, uu_fcqs)
mcorr.fcqs.vec
```
Note that the results range from $0$ to $1$, and a number closer to $d$ indicates better performance.

To better understand the output, we can visualize the estimated functional central quantile subspace directions and the relationship between the true and estimated predictors.

First, we can plot the relationship between the response variable $Y$ and the first sufficient predictor
```{r, fig.align = "center"}
library(ggplot2)

# Create a dataframe for plotting
df_y_vs_score <- data.frame(score = mfpca.scores[, 1], y = y)

# Scatter plot
ggplot(df_y_vs_score, aes(x = score, y = y)) +
  geom_point() +
  geom_smooth(method = "gam", se = FALSE, color = "red") +
  labs(x = "First sufficient predictor", y = "y") +
  theme_minimal()
```

Then we can compare the estimated inner product from the `fcqs` function to the true inner products used in data generation. 
```{r, fig.align = "center"}
# Create a dataframe for plotting
df_inner_products <- data.frame(estimated = output1$betax, true = uu_fcqs)

# Scatter plot
ggplot(df_inner_products, aes(x = true, y = estimated)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  labs(x = "True Inner Products", y = "Estimated Inner Products") +
  theme_minimal()
```

Note that the reference line (slope = 1) indicates perfect agreement. The overall positive trend and the clustering around the reference line suggest that the `fcqs` function reasonably accurately estimates the inner products.

# References
1. Cardot, H., Crambes, C., and Sarda, P. (2005). Quantile regression when the covariates are functions. *Journal of Nonparametric Statistics*, 17, 841--856.

1. Chen, K., and M\"uller, H.-G. (2012). Conditional quantile analysis when covariates are functions, with application to growth data. *Journal of the Royal Statistical Society. Series B (Statistical Methodology)*, 74(1), 67--89. 

1. Christou, E., Solea, E., Wang, S., and Song, J. (2024+). Sufficient dimension reduction for conditional quantiles for functional data. Under review. 

1. Ferr\'e, L., and Yao, F. (2003). Functional sliced inverse regression analysis. *Statistics*, 37, 475--488.

1. Guerre, E., and Sabbah, C. (2012). Uniform bias study and Bahadur representation for local polynomial estimators of the conditional quantile function. *Econometric Theory*, 28(1), 87--129.  

1. Kato, K. (2012). Estimation in functional linear quantile regression. *The Annals of Statistics*, 40: 3108--3136.

1. Koenker, R., and Bassett, G. (1978) Regression quantiles. *Econometrica*, 46, 33-50

1. Lian, H., and Li, G. (2014). Series expansion for functional sufficient dimension reduction. *Journal of Multivariate analysis*, 124, 150--165. 

1. Li, B, and Song, J. (2022). Dimension reduction for functional data based on weak conditional moments. *The Annals of Statistics*, 50(1), 107--128. 

1. Ramsay, J. O., and Dalzell, C. J. (1991). Some tools for functional data analysis, *Journal of the Royal Statistical Society: Series B (Methodological)*, 53, 539--572.

1. Wang, G., Lin, N., and Zhang, B. (2013). Functional contour regression. *Journal of Multivariate Analysis*, 116, 1--13. 
